Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2017-10-16 TACAS2018/paper.tex
sec 4 edit (partial)

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Model semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of Continuous time systems}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 The main problem with the above discretization is that is does not allow us to observe the system at arbitrary points in time, only at integer multiples of the sample time. While this is adequate for the controller, it is not sufficient for verification. We must therefore establish a new model that allows us to observe the system at a given time $t$.
 %We achieve this by using acceleration as described below.

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelrated dynamics using reals for systems without inputs}\label{sec:real_discrete_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 The state $\vec{x}_T$ present at time $T$ can be expressed as a linear function of $\vec{x}_0$ as described below
 \begin{theorem}
 Given $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, the expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  $\vec{x}_T$ is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0 : k \in (0,\infty)$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
% \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics using reals for systems with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0+(\mat{I}-\mat{A}_d^k\mat{B}\vec{u}) : k \in (0,\infty)$ is a witness of the system at time $kT_d$.
 \end{corollary}
% \begin{proof}
% We begin by expanding the equation
% \begin{align}
% \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\nonumber\\
% &=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u}\nonumber\\
% &= \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}
% \end{align}
% where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
% \end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics using reals for systems with discrete time feedback inputs}\label{sec:real_discrete_feedback_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous time systems}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Note that the discretization in Sec.~\ref{sec:ssrepresentation} only 
 allows us to observe the system at integer multiples of the sample time.
 As we are interested in ensuring that certain properties of the system hold at any given
 time (not just integer multiples of the sample time),
 we accelerate the dynamics of the system (according to the model semantics in Sec.~\ref{sec:model_semantics}) in order to compute the state $\vec{x}_T$  at a given time $T$.

 %While this is adequate for the controller, it is not sufficient for verification. We must therefore establish a new model that allows us to observe the system at a given time $t$.
 %We achieve this by using acceleration as described below.

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for systems without inputs}\label{sec:real_discrete_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Given a specific time $T$, the state $\vec{x}_T$ present at $T$ can be expressed as a linear function of $\vec{x}_0$ as described below. Note that, as the system has no inputs, its dynamics are described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$.

 \begin{theorem}
 Given $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, we have
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \text{where }\mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{and } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  We call $\vec{x}_T$ a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 %% We can apply Th.~\ref{eq:continuous_tube_dyn} to discretized systems without inputs in order to
 %% obtain a witness ..

\todo{Can the corallary be motivated? The general motivation of the section was observing the system at times that are not necessarily integer multiple of the sampling time...} 
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0 : k \in (0,\infty)$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
% \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for systems with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in observing systems with parametric
 inputs, described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$.
 For such systems, a witness $\vec{x}_T$ is computed next.
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
\end{theorem}
\todo{\eqref{eq:continuous_tube_dyn} is in the appendix}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x}_0+(\mat{I}-\mat{A}_d^k\mat{B}\vec{u}) : k \in (0,\infty)$ is a witness of the system at time $kT_d$.
 \end{corollary}
% \begin{proof}
% We begin by expanding the equation
% \begin{align}
% \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\nonumber\\
% &=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u}\nonumber\\
% &= \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}
% \end{align}
% where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
% \end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for systems with discrete time feedback inputs}\label{sec:real_discrete_feedback_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
